---
title: "Exploring the gapminder"
author: "Jue Zhou"
date: "11/21/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      error = FALSE)
library(here)
library(gapminder)
library(geonames)
library(tidyverse)
library(countrycode)
library(dplyr)

```

## Gapminder

Here I use the `gapminder` data to explore the relationship between population density and life expectency.

The population density is calculated using the information of 'area in squre kms' from `geonames` package.
```{r countryInfo}
glimpse(gapminder)
# access country information via geonames API
countryInfo <- GNcountryInfo()
glimpse(countryInfo)
```

```{r merge}
# recode the country name in gapminder to match it with countryInfo
gapminder_new <- gapminder %>% 
  mutate(iso3c = countrycode(country, origin = 'country.name', destination = 'iso3c')) %>% 
  # merge countryInfo into gapminder
  left_join(countryInfo, by = c("iso3c" = "isoAlpha3")) %>% 
  # select the columns to use
  select(country, year, lifeExp, pop, areaInSqKm, continentName) %>% 
  # calculate the population density for each observation
  mutate(areaInSqKm = as.numeric(areaInSqKm)) %>% 
  mutate(popden = pop/areaInSqKm) %>% 
  rename(Continent = "continentName")
```


```{r graph1}
# Produce an updated graph using population density
gapminder_new %>%
  ggplot(aes(log(popden), lifeExp)) +
  geom_point(alpha = 0.1) +
  geom_smooth() +
  labs(x = "Population density(logged) per square km",
       y = "Average Life Expectancy",
       title = "Relationship between Life Expectancy and Population Density") +
  theme_minimal()
```

This graph above shows the relationship between population density and life expectancy. Generally, as population density increases, life expectancy increases.

```{r graph2}
# Compare the relationship across continents
gapminder_new %>%
  group_by(Continent) %>%
  ggplot(aes(log(popden), lifeExp, color = Continent)) +
  geom_point(alpha = 0.1) +
  geom_smooth(se = FALSE, method = "loess") +
  labs(x = "Population density(logged) per square km",
       y = "Average Life Expectancy",
       title = "Relationship between Life Expectancy and Population Density") +
  theme_minimal()
  
```

The second graph shows the relationship between population density and life expectancy across continents. For Africa, Asia and South America, the positive relationship between population density and life expectancy sustains, which indicates this relationship might mainly apply to developing countries.

```{r, include=FALSE}
devtools::session_info()
```
